---
title: "DS311 - R Lab Assignment"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
# rows = observations, I think
# columns = variables

# Answer:
dim(mtcars)
print(paste("There are total of ", ncol(mtcars)," variables and ", nrow(mtcars)," observations in this data set."))
```

b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
# str(mtcars)
summary(mtcars)

# Answer:
# Unfortunately, I can't tell what the abbreviations stand for.  But,
# from the values, it looks to me like the following are discrete: 
# cyl, vs, am, gear and carb.  The following are continuous: 
# mpg, disp and hp (even though they are whole numbers), drat, wt and qsec.
print("There are 5 discrete variables and 6 continuous variables in this data set.")
```

c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!

v <- var(mtcars$mpg)
s <- sd(mtcars$mpg)
m <- mean(mtcars$mpg)

print(paste("The average of Mile Per Gallon from this data set is ", m , " with variance ", v , " and standard deviation", s, "."))
```

d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
# mtcars$cyl
# x <- aggregate(mtcars::group_by(mtcars, cyl))
# x <- tbl(mtcars$cyl == 4)
# head(fourcyl)
# mean(x$mpg)
tapply(mtcars$mpg, mtcars$cyl, summary)
# x <- data.frame(group_by(mtcars, cyl))
# x
```

e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
# install.packages("tidyverse")
# library(tidyverse)
mtcars %>% select(cyl, gear) %>% table()

print("The most common car type in this data set is car with 8 cylinders and 3 gears. There are total of 12 cars belong to this specification in the data set.")
```

*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
ggplot(PlantGrowth, aes(x = group, y = weight, fill = group)) +
    geom_boxplot() +
    geom_jitter(shape = 15,
        color = "steelblue",
        position = position_jitter(0.21)) +
   theme_minimal()

```

Result:

=> Report a paragraph to summarize your findings from the plot!

There are three groups. The first group is the control group.  The second group are plants that underwent treatment 1.  The third group is the remaining plants, who underwent treatment 2.  This boxplot shows that most of the control group is between weights 4.5 to 5.6 in the middle.  But the plants in treatment 1 tended to weigh less, with most between 4.0 and 5.0.  Treatment 2 plants are the opposite, with all the plants weighing more than 5.0, except for one.

b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
ggplot(data=mtcars, aes(x=mpg)) + geom_histogram()


print("Most of the cars in this data set are in the class of 16 miles per gallon.")
```


c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)
# USArrests

# Enter your code here!
pairs(USArrests)
pairs(USArrests[,1:2])

```

Result:

=> Report a paragraph to summarize your findings from the plot!

I have found that murder and assault are positively correlated, so that when one goes up the other generally does also.  The number of assaults is generally about 23 times the number of murders.  As a state's population increases, so does the murder count and assault count.

***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
housingData %>% group_by(Neighborhood) %>% summarise(mean(Market.Value.per.SqFt))

```

b. Create multiple plots to demonstrate the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
# Can't use pairs() or heatmap(), because some of the data is not numeric.
# head(housingData)
# heatmap(housingData)
# install.packages("ggplot2")
library(ggplot2)
ggplot(housingData, aes(x=Market.Value.per.SqFt, y=Boro)) + geom_boxplot() + scale_y_discrete(limits=unique((housingData$Boro)))
# %>% group_by(Neighborhood)
# %>% summarise(mean(Market.Value.per.SqFt))

```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!
Manhattan clearly dominates the most expensive market values per square foot.  Out of the 1290 most expensive housing units, 1207 of them are in Manhattan! That's 94%. But what's interesting is that Manhattan also has a few housing units in the least expensive end.  The horizontal line that represents Manhattan is very long, whereas the horizontal line representing Staten Island is very skinny.  So, all the housing on Staten Island is inexpensive, compared to the other boroughs.  <-- That's the correct spelling, not "boro".

-- Steve Stilson
